Evolution of ORM Systems: Formal Foundations

### 2013-01-14 doc/description.tex
change reference of a row
getAssocByname, propertyByName

@@

\paragraph{Insert Value} The function inserts a new pair into an existing row.
\begin{align*}
	\mathbf{insertValues} : Row \times Pair* \times Database \rightarrow Daatabase \\
	insertValues(r, p, d) = \begin{cases}
 		\inference{|p| > 1}{insertValues(r, head(p), d)\bullet insertValues(r, tail(p), d)}  \\ \\
 		\inference{|p| = 1 \wedge \forall e \in pairs(r) : reference(e) \neq reference(p) \\ \wedge \exists t \in tables(d) : name(t) = refernces(r), \\ \wedge \exists c \in columns(t) : reference(p) = name(c)}{pairs(r) \cup p} \\ \\
	 	 Inconsistent \; d
 		\end{cases} \\
r \lhd Row, p = Pair*,d \lhd Databases
\end{align*}
\paragraph{Copy Value} The function copy values from one column in a table to another column in the same or in another table - according to mapping function $f_m$. 
\begin{align*}
	\mathbf{copyValue} : Label \times Label \times Label \times (Row \times Row) \times Database \rightarrow Database \\
	copyValue(l_s, l_t, l_{sc}, l_{tc} f_m, d) = \begin{cases}
		\inference{\exists t_s \in tables(d) : name(t_s) = l_s, \\ t_t \in tables(d) : name(t_t) = l_t, \\ c \in columns(t_s) : name(c) = l_{sc}, \\ e \in columns(t_t) : name(e) = l_{tc}}{\begin{gathered}
			insertValues(second(p), \\ [l_{tc},value(pairByRef(l_{sc}, first(p)))],d)
		\end{gathered}}
 	\end{cases}
\end{align*}

\paragraph{Copy rows}
\begin{align*}
	\mathbf{copyRows} : Label \times Label \times (Row \times Row) \times Database \rightarrow Database \\
	copyRows(l_s, l_t, f_m, d) = \begin{cases}
 	\inference{|f_m| = \varnothing}{d} \\\\
 	\inference{|f_m| = 1 \\ \wedge \exists t_s, t_t \in tables(d) : name(t_s) = l_s, name(t_t) = l_t \\ \wedge t_s \cong t_t }{data(d) \cup insertRow(changeRef(l_t, first(f_m)), d)} \\\\
 	\inference{|f_m| > 1}{\begin{gathered}
		 d \bullet copyRows(l_s, l_t, head(f_m)) \bullet copyRows(l_s, l_t, tail(f_m))
	\end{gathered}}
  \end{cases}
\end{align*}

@@

\paragraph{Rename Association}
\begin{align*}
RenameAssociation : Label \times Label \times Label \times State \rightarrow State \\
\Psi(RenameAssocaiation(l_c, l_o, l_n), s) = application(s) \circ renAssoc(l_c, l_o, l_n)\\
\Phi(RenameAssocaiation(l_c, l_o, l_n), s) = \begin{cases}
	\inference{\exists a  \in getAssocByname(application(s)) : \\
  name(a) = l_c \\ \wedge startCardinality(a) > 1 }{database(s) \bullet alterTable(l_c, l_o, l_n)}
 \\ \\
  \inference{\exists a  \in getAssocByname(application(s)) : \\
  name(a) = l_c \\ \wedge startCardinality(a) \leq 1 }{\begin{gathered}
database(s) \\ \bullet alterForeignKey(endClass(a) ,l_c, l_o, l_n)
\end{gathered}}
 \end{cases}
\\\\
%	alterAssoc(l_c, l_o, l_n, d) = \begin{cases}
%?? \inference{\exists t \in tables(a) : name(t) = l_c \\ \wedge \exists f \in fks(t) : name(c) = l_o \\ \wedge \forall e \in fks(t) : name(e) \neq l_n}{\begin{gathered}
%f.name = l_n 
%\end{gathered}}
%\\ \\
%\inference{\exists t \in tables(a) : name(t) = l_o \\ \wedge \forall e \in tables(t) : name(e) \neq l_n}{\begin{gathered}
%t.name = l_n 
%\end{gathered}}
% \end{cases}
% \\
l_c, l_o, l_n \in Label, a \lhd Applications, s \in Sates, d \lhd Database
\end{align*}

>>>>

\paragraph{Get Property or Association}
\begin{align*}
	\mathbf{propertyByName} : Label \times Label \times Application \rightarrow Property \\
	propertyByName(l_c, l_p, a) = \inference{\exists c \in classes(a): name(c) = l_c \\ \wedge \exists p \in properties(c) : name(p) = l_p}{p} \\
	l_c, l_p \in Label, a \lhd Application
\end{align*}
\begin{align*}
	\mathbf{associationByName} : Label \times Label \times Application \rightarrow Association \\
	associationByName(l_c, l_a, a) = \inference{\exists c \in classes(a): name(c) = l_c \\ \wedge \exists o \in associations(c) : name(o) = l_a}{o} \\
	l_c, l_a \in Label, a \lhd Application
\end{align*}

@@

\paragraph{Insert Value} The function inserts a new pair into an existing row.
\begin{align*}
	\mathbf{insertValues} : Row \times Pair* \times Database \rightarrow Daatabase \\
	insertValues(r, p, d) = \begin{cases}
 		\inference{|p| > 1}{insertValues(r, head(p), d)\bullet insertValues(r, tail(p), d)}  \\ \\
 		\inference{|p| = 1 \wedge \forall e \in pairs(r) : reference(e) \neq reference(p) \\ \wedge \exists t \in tables(d) : name(t) = refernces(r), \\ \wedge \exists c \in columns(t) : reference(p) = name(c)}{pairs(r) \cup p} \\ \\
	 	 Inconsistent \; d
 		\end{cases} \\
r \lhd Row, p = Pair*,d \lhd Databases
\end{align*}
\paragraph{Copy Value} The function copy values from one column in a table to another column in the same or in another table - according to mapping function $f_m$. 
\begin{align*}
	\mathbf{copyValue} : Label \times Label \times Label \times (Row \times Row) \times Database \rightarrow Database \\
	copyValue(l_s, l_t, l_{sc}, l_{tc} f_m, d) = \begin{cases}
		\inference{\exists t_s \in tables(d) : name(t_s) = l_s, \\ t_t \in tables(d) : name(t_t) = l_t, \\ c \in columns(t_s) : name(c) = l_{sc}, \\ e \in columns(t_t) : name(e) = l_{tc}}{\begin{gathered}
			insertValues(second(p), \\ [l_{tc},value(pairByRef(l_{sc}, first(p)))],d)
		\end{gathered}}
 	\end{cases} \\
 	l_s, l_t, l_{sc}, l_{tc} \in Label, d \lhd Database
\end{align*}

\paragraph{Copy rows}
\begin{align*}
	\mathbf{copyRows} : Label \times Label \times (Row \times Row) \times Database \rightarrow Database \\
	copyRows(l_s, l_t, f_m, d) = \begin{cases}
 	\inference{|f_m| = \varnothing}{d} \\\\
 	\inference{|f_m| = 1 \\ \wedge \exists t_s, t_t \in tables(d) : name(t_s) = l_s, name(t_t) = l_t \\ \wedge t_s \cong t_t }{data(d) \cup insertRow(changeRef(l_t, first(f_m)), d)} \\\\
 	\inference{|f_m| > 1}{\begin{gathered}
		 d \bullet copyRows(l_s, l_t, head(f_m)) \bullet copyRows(l_s, l_t, tail(f_m))
	\end{gathered}}
  \end{cases} \\
  l_s, l_t \in Label, d \lhd Database
\end{align*}

\paragraph{Change ref}
\begin{align*}
	\mathbf{changeRef} : Label \times Row \rightarrow Row \\
	changeRef(l, r) = \inference{}{reference(r) = l } \\
	l \in Label, r \lhd Row
\end{align*}

@@

\paragraph{Rename Association}
\begin{align*}
RenameAssociation : Label \times Label \times Label \times State \rightarrow State \\
\Psi(RenameAssocaiation(l_c, l_o, l_n), s) = application(s) \circ renAssoc(l_c, l_o, l_n)\\
\Phi(RenameAssocaiation(l_c, l_o, l_n), s) = \begin{cases}
	\inference{\exists a  \in associationByName(application(s)) : \\
  name(a) = l_c \\ \wedge startCardinality(a) > 1 }{database(s) \bullet alterTable(l_c, l_o, l_n)}
 \\ \\
  \inference{\exists a  \in associationByName(application(s)) : \\
  name(a) = l_c \\ \wedge startCardinality(a) \leq 1 }{\begin{gathered}
database(s) \\ \bullet alterForeignKey(endClass(a) ,l_c, l_o, l_n)
\end{gathered}}
 \end{cases}
\\\\
%	alterAssoc(l_c, l_o, l_n, d) = \begin{cases}
%?? \inference{\exists t \in tables(a) : name(t) = l_c \\ \wedge \exists f \in fks(t) : name(c) = l_o \\ \wedge \forall e \in fks(t) : name(e) \neq l_n}{\begin{gathered}
%f.name = l_n 
%\end{gathered}}
%\\ \\
%\inference{\exists t \in tables(a) : name(t) = l_o \\ \wedge \forall e \in tables(t) : name(e) \neq l_n}{\begin{gathered}
%t.name = l_n 
%\end{gathered}}
% \end{cases}
% \\
l_c, l_o, l_n \in Label, a \lhd Applications, s \in Sates, d \lhd Database
\end{align*}
